Graphical Tests for Power Comparison of Competing Designs

### 2012-03-29 setup.tex
small wording changes
second table

The perceptual tasks involved in decoding the designs fit into the general framework of the piechart versus barchart discussion. However, we are not dealing with assessing angles, which is known to be a perceptual harder task than assessments of heights or widths in bars, see e.g. \cite{cleveland:1984, robbins:2004, few:2009}. Instead, the perceptual tasks consists of comparisons along a common axis (in the barcharts) and a common origin (in the polar charts). The difference in designs is therefore based on how well we are able to judge deviations from a horizontal line compared to deviations from a circle. Based on this, the charts with added reference lines provide us with exactly the frame we compare with and should, therefore, be the `better' designs - either in speed or accuracy.

@@

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
%\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrrrl}
  \hline
& & Estimate & Error & $z$-value & $p$-value &\\ 
  \hline
\bf design & euclid & -0.08 & 0.39 & -0.21 & 0.84 &\\ 
&polar & -1.98 & 0.32 & -6.13 & 0.00 & ***\\ [3pt]
\multicolumn{2}{l}{\bf main effects} &&&&&\\
& reference line  & -0.14 & 0.26 & -0.53 & 0.59 & \\ [1pt]
&  sample size & 0.27 & 0.04 & 6.31 & 0.00 & ***\\[1pt]
 &offset:  90 degrees& -0.43 & 0.37 & -1.18 & 0.24 &\\ 
  & 180 degrees& -0.89 & 0.35 & -2.51 & 0.01 & **\\ 
  & 270 degrees& 0.21 & 0.38 & 0.55 & 0.58 &\\ [3pt]
\multicolumn{2}{l}{\bf interactions} &&&&&\\
&  polar:line & 0.51 & 0.35 & 1.44 & 0.15 &\\ [1pt]
&    polar:sample size & -0.23 & 0.05 & -5.02 & 0.00 & ***\\[1pt]
&    polar:offset 90 & 0.64 & 0.49 & 1.30 & 0.20 \\ 
&  polar:offset 180 & 0.91 & 0.47 & 1.92 & 0.05 & .\\ 
&    polar:offset 270 & -0.73 & 0.54 & -1.35 & 0.18 &\\
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}
}
\end{center}
\caption{\label{tbl:correct} Output of a generalized linear mixed effects model for power of lineups (i.e. probability of identifying the data plot) for comparing designs. Included are two-way interactions with sample size and shifts in wind direction (offset). Results are based on  976 lineup evaluations by 115 participants. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std..Error & t.value & pval \\ 
  \hline
(Intercept) & 3.58 & 0.09 & 41.34 & 0.00 \\ 
  polar & 0.22 & 0.10 & 2.17 & 0.03 \\ 
  reflineTRUE & 0.08 & 0.06 & 1.50 & 0.13 \\ 
  sample\_size & -0.03 & 0.00 & -8.73 & 0.00 \\ 
  offset90 & -0.06 & 0.08 & -0.72 & 0.47 \\ 
  offset180 & 0.16 & 0.08 & 2.07 & 0.04 \\ 
  offset270 & -0.04 & 0.07 & -0.54 & 0.59 \\ 
  polar:reflineTRUE & -0.03 & 0.08 & -0.40 & 0.69 \\ 
  polar:sample\_size & 0.03 & 0.01 & 6.25 & 0.00 \\ 
  polar:offset90 & 0.14 & 0.11 & 1.23 & 0.22 \\ 
  polar:offset180 & -0.17 & 0.11 & -1.55 & 0.12 \\ 
  polar:offset270 & 0.17 & 0.11 & 1.50 & 0.13 \\ 
   \hline
\end{tabular}}
\end{center}
\caption{\label{tbl:time} Model output for (log) time taken. }
\end{table}

>>>>

The perceptual tasks involved in decoding the designs fit into the general framework of the piechart versus barchart discussion. However, we are not dealing with assessing angles, which is known to be a perceptual harder task than assessments of heights or widths in bars, see e.g. \cite{cleveland:1984, robbins:2004, Kosslyn:2006, few:2009}. Instead, the perceptual tasks consists of comparisons along a common axis (in the barcharts) and a common origin (in the polar charts). The difference in designs is therefore based on how well we are able to judge deviations from a horizontal line compared to deviations from a circle. Based on this, the charts with added reference lines provide us with exactly the frame we compare with and should, therefore, be the `better' designs - either in speed or accuracy.

@@

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
%\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrrrl}
  \hline
& & Estimate & Error & $z$-value & $p$-value &\\ 
  \hline
\bf design & euclid & -0.08 & 0.39 & -0.21 & 0.84 &\\ 
&polar & -1.98 & 0.32 & -6.13 & 0.00 & ***\\ [3pt]
\multicolumn{2}{l}{\bf main effects} &&&&&\\
& reference line  & -0.14 & 0.26 & -0.53 & 0.59 & \\ [1pt]
&  sample size & 0.27 & 0.04 & 6.31 & 0.00 & ***\\[1pt]
 &offset:  90 degrees& -0.43 & 0.37 & -1.18 & 0.24 &\\ 
  & 180 degrees& -0.89 & 0.35 & -2.51 & 0.01 & **\\ 
  & 270 degrees& 0.21 & 0.38 & 0.55 & 0.58 &\\ [3pt]
\multicolumn{2}{l}{\bf interactions} &&&&&\\
&  polar:line & 0.51 & 0.35 & 1.44 & 0.15 &\\ [1pt]
&    polar:sample size & -0.23 & 0.05 & -5.02 & 0.00 & ***\\[1pt]
&    polar:offset 90 & 0.64 & 0.49 & 1.30 & 0.20 \\ 
&  polar:offset 180 & 0.91 & 0.47 & 1.92 & 0.05 & .\\ 
&    polar:offset 270 & -0.73 & 0.54 & -1.35 & 0.18 &\\
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}
}
\end{center}
\caption{\label{tbl:correct} Output of a generalized linear mixed effects model for power of lineups (i.e. probability of identifying the data plot) for comparing designs. Included are two-way interactions with sample size and shifts in wind direction (offset). Euclidean designs without a reference line at offset 0 are used as 
 Results are based on  976 lineup evaluations by 115 participants. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
%\rowcolors{2}{white}{lightgray}
\begin{tabular}{lrrrrrl}
  \hline
& & Estimate & Error & $t$-value & \multicolumn{2}{l}{approx $p$-value} \\ 
  \hline
\bf design & Intercept & 3.58 & 0.09 & 41.34 & 0.00 & ***\\ 
&  polar & 0.22 & 0.10 & 2.17 & 0.03 & * \\ [2pt]
\multicolumn{2}{l}{\bf covariates}\\
&reference line & 0.08 & 0.06 & 1.50 & 0.13 \\ [1pt]
 & sample size & -0.03 & 0.00 & -8.73 & 0.00 & ***\\ [1pt]
&  offset: 90 & -0.06 & 0.08 & -0.72 & 0.47 \\ 
& 180 & 0.16 & 0.08 & 2.07 & 0.04 & *\\ 
&  270 & -0.04 & 0.07 & -0.54 & 0.59 \\ [2pt]
\multicolumn{2}{l}{\bf interactions}\\
&  polar:line & -0.03 & 0.08 & -0.40 & 0.69 \\ [1pt]
&    polar:sample size & 0.03 & 0.01 & 6.25 & 0.00 & ***\\ [1pt]
&    polar:offset 90 & 0.14 & 0.11 & 1.23 & 0.22 \\ 
&    polar:offset 180 & -0.17 & 0.11 & -1.55 & 0.12 \\ 
&    polar:offset 270 & 0.17 & 0.11 & 1.50 & 0.13 \\ 
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}}
\end{center}
\caption{\label{tbl:time} Model output for linear mixed effects model of (log) time taken, barchart design is baseline. }
\end{table}
